---
title: "Setting Up the RRC R Tools Plugin for Claude Code"
author: "Cathy Thomas"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    df-print: kable
    theme: cosmo
    embed-resources: true
execute:
  echo: false
  warning: false
  message: false
---

## Goal

This guide walks through installing the `rrc-R-tools` Claude Code plugin, which gives you a shared set of R and Quarto skills that load automatically in every Claude Code session.

This is a Claude Code-specific setup. 

## What a Plugin Is

Claude Code can be extended with **plugins**: packages of skills, commands, and other behaviors that load automatically whenever you start a session. Skills are context documents that shape how Claude responds. The `r-style` skill, for example, tells Claude to prefer tidyverse idioms and `here::here()` for file paths without you having to say so every time.

Plugins are the team mechanism for sharing skills. Rather than each person maintaining their own copy of the R style guide in `~/.claude/skills/`, we maintain one shared version in a plugin repo on GitHub. When the skill is updated, everyone gets the new version the next time Claude Code refreshes its plugin cache.

## Plugins vs Personal Skills

You can also keep skills that are just for you. Personal skills live in `~/.claude/skills/` on your own machine and are not shared with anyone. A personal skill might capture your preferred writing style, your go-to packages for a specific project type, or any other preference that doesn't need to be team-wide. To add one, create a subfolder in `~/.claude/skills/` with a `SKILL.md` file inside it.

You do not need to write the skill file yourself. You can ask Claude to create one for you ("create a personal skill that..."), or use Claude's built-in skill-creator skill, which guides you through building, testing, and refining a new skill.

Plugin skills and personal skills coexist. Both load at the start of every session.

## Skills Included in the RRC R Tools Plugin

| Skill | What it does |
|---|---|
| `r-style` | Applies tidyverse style when writing or reviewing R code: pipe-based workflows, dplyr/tidyr/ggplot2, `here::here()` for repo paths. Loads on any `.R`, `.Rmd`, or `.qmd` work without prompting. |
| `diagnose-before-fixing` | When you share an error, Claude explains what it means and identifies the root cause before proposing a fix, then waits for your confirmation before making edits. |
| `quarto-render-sync` | On-demand skill for setting up a post-render hook that renames HTML output and copies it to a network drive (Egnyte / Z: drive). |
| `quarto-scaffold` | On-demand skill for scaffolding a new Quarto project with the right folder structure, `_quarto.yml`, template `.qmd`, setup chunk, and `.gitignore`. |

The first two skills (`r-style` and `diagnose-before-fixing`) load automatically on every session. The last two (`quarto-render-sync` and `quarto-scaffold`) are reference skills. To invoke them, describe what you want in plain language ("scaffold a new Quarto project" or "set up a render-and-sync workflow"), or use a slash command directly: `/quarto-scaffold` or `/quarto-render-sync`. This works in both the Claude Code terminal and the Claude Code GUI in the Claude App desktop.

## Prerequisites

### Claude desktop App

The Claude desktop App is where you do your day-to-day work. If you don't have it yet, download it from [claude.ai](https://claude.ai/download) and log in.

### Claude Code CLI

The plugin install commands require the **Claude Code CLI** — a separate tool from the desktop App. Having the App does not give you the `claude` terminal command; you need to install the CLI separately. You only need it once for this setup.

**Install the CLI.** Open PowerShell and run:

```{r}
#| echo: true
#| eval: false
irm https://claude.ai/install.ps1 | iex
```

**Fix PATH if needed.** The installer may warn that the install location is not in your PATH. If so, run this before closing PowerShell (no need to change anything — it uses your current username automatically):

```{r}
#| echo: true
#| eval: false
[Environment]::SetEnvironmentVariable("PATH", $env:PATH + ";$env:USERPROFILE\.local\bin", "User")
```

Then close and reopen PowerShell.

**Verify.** Run `claude --version`. You should see a version number. If you still get "not recognized," close and reopen PowerShell once more.

### GitHub access

You need access to the `RRC-Associates/claude-rrc-plugins` GitHub repository. No local clone is needed — Claude Code fetches the plugin directly from GitHub.

## Setup

::: callout-important
**Plugin management is CLI-only.** All of the `/plugin ...` commands and `/reload-plugins` run inside a Claude Code **CLI session** (the `claude` command in PowerShell), not in the desktop App GUI. The GUI loads and uses the skills, but it doesn't expose the plugin-management slash commands. You only need to do these setup steps once; afterward, the skills are available in both CLI and GUI sessions.
:::

### Step 1: Open a Claude Code session

In PowerShell, run:

```{r}
#| echo: true
#| eval: false
claude
```

You will be prompted to:

1. **Sign in** — select the option to log in with your Anthropic subscription. A verification code will be displayed; paste it back into the session when prompted.
2. **Pick a mode** — select the default mode to continue.

Once signed in, you'll see the `>` prompt. You're now inside Claude Code.

### Step 2: Add the marketplace

At the `>` prompt, run:

```{r}
#| echo: true
#| eval: false
/plugin marketplace add RRC-Associates/claude-rrc-plugins
```

This registers the RRC plugin repository as a source Claude Code can install from. You only need to do this once.

### Step 3: Install the plugin

Still at the `>` prompt, run:

```{r}
#| echo: true
#| eval: false
/plugin install rrc-R-tools@claude-rrc-plugins
```

Claude Code will show a confirmation screen with the plugin name and description. Select **Install for you (user scope)** and press Enter.

User scope installs the plugin for your account across all projects. Do not choose project scope unless you have a specific reason.

### Step 4: Enable auto-update

By default, Claude Code only auto-checks for updates from Anthropic's own marketplaces — third-party marketplaces like `claude-rrc-plugins` have auto-update disabled out of the box. Turn it on now so you don't miss future skill updates:

1. At the `>` prompt, run `/plugin`.
2. Open the **Marketplaces** tab and select `claude-rrc-plugins`.
3. Choose **Enable auto-update**.

With this enabled, Claude Code checks for new versions at session startup and prompts you to run `/reload-plugins` whenever an update is available.

### Step 5: Verify

Close and reopen Claude Code (CLI or App GUI), then start a new session. Ask Claude: "What skills do you have available?" You should see `r-style` and `diagnose-before-fixing` in the response. `quarto-render-sync`, `quarto-scaffold`, and `quarto-word` are on-demand skills — they won't appear in the auto-load list but are available when you need them (just describe what you want or use the slash commands directly, e.g. `/quarto-scaffold`, `/quarto-render-sync`, `/quarto-word`).

## Keeping the Plugin Current

You enabled auto-update during setup (Step 4), so Claude Code checks for new versions at the start of each session.

### From the CLI

When an update is available, Claude Code prompts you to run `/reload-plugins` — that loads the new versions into your active session. That's the entire flow.

**To force a refresh manually** (e.g., a change just landed and you don't want to wait for the next session start):

```
/plugin marketplace update claude-rrc-plugins
/reload-plugins
```

The first command pulls the latest catalog and plugin versions from the GitHub repo; the second loads them into your session.

### From the desktop App GUI

The `/plugin ...` family and `/reload-plugins` are **not available** in the GUI input — they're CLI-only. To pick up an update from the GUI, **fully quit the app and relaunch it**.

::: callout-important
On Windows, closing the window ("X-ing out") doesn't always end the process. Open **Task Manager**, end the `Claude` process (and any background helper processes), then relaunch the app. After this restart, the new skill version will be loaded.
:::

If you want to force a marketplace refresh before relaunching the GUI, open a CLI session (`claude` in PowerShell), run the two-command pair above, then fully quit and relaunch the GUI.

### Suggesting changes

To suggest a change to one of the skills, reach out to Cathy. Changes go into the GitHub repo and propagate to everyone on the next update.
