---
title: "Setting Up the RRC R Tools Plugin for Claude Code"
author: "Cathy Thomas"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    df-print: kable
    theme: cosmo
    embed-resources: true
execute:
  echo: false
  warning: false
  message: false
---

## Goal

This guide walks through installing the `rrc-R-tools` Claude Code plugin, which gives you a shared set of R and Quarto skills that load automatically in every Claude Code session.

This is a Claude Code-specific setup. 

## What a Plugin Is

Claude Code can be extended with **plugins**: packages of skills, commands, and other behaviors that load automatically whenever you start a session. Skills are context documents that shape how Claude responds. The `r-style` skill, for example, tells Claude to prefer tidyverse idioms and `here::here()` for file paths without you having to say so every time.

Plugins are the team mechanism for sharing skills. Rather than each person maintaining their own copy of the R style guide in `~/.claude/skills/`, we maintain one shared version in a plugin repo on GitHub. When the skill is updated, everyone gets the new version the next time Claude Code refreshes its plugin cache.

## Plugins vs Personal Skills

You can also keep skills that are just for you. Personal skills live in `~/.claude/skills/` on your own machine and are not shared with anyone. A personal skill might capture your preferred writing style, your go-to packages for a specific project type, or any other preference that doesn't need to be team-wide. To add one, create a subfolder in `~/.claude/skills/` with a `SKILL.md` file inside it.

You do not need to write the skill file yourself. You can ask Claude to create one for you ("create a personal skill that..."), or use Claude's built-in skill-creator skill, which guides you through building, testing, and refining a new skill.

Plugin skills and personal skills coexist. Both load at the start of every session.

## Skills Included in the RRC R Tools Plugin

| Skill | What it does |
|---|---|
| `r-style` | Applies tidyverse style when writing or reviewing R code: pipe-based workflows, dplyr/tidyr/ggplot2, `here::here()` for repo paths. Loads on any `.R`, `.Rmd`, or `.qmd` work without prompting. |
| `diagnose-before-fixing` | When you share an error, Claude explains what it means and identifies the root cause before proposing a fix, then waits for your confirmation before making edits. |
| `quarto-render-sync` | On-demand skill for setting up a post-render hook that renames HTML output and copies it to a network drive (Egnyte / Z: drive). |
| `quarto-scaffold` | On-demand skill for scaffolding a new Quarto project with the right folder structure, `_quarto.yml`, template `.qmd`, setup chunk, and `.gitignore`. |

The first two skills (`r-style` and `diagnose-before-fixing`) load automatically on every session. The last two (`quarto-render-sync` and `quarto-scaffold`) are reference skills. To invoke them, describe what you want in plain language ("scaffold a new Quarto project" or "set up a render-and-sync workflow"), or use a slash command directly: `/quarto-scaffold` or `/quarto-render-sync`. This works in both the Claude Code terminal and the Claude Code GUI in the Claude App desktop.

## Prerequisites

### Claude desktop App

The Claude desktop App is where you do your day-to-day work. If you don't have it yet, download it from [claude.ai](https://claude.ai/download) and log in.

### Claude Code CLI

The plugin install commands require the **Claude Code CLI** — a separate tool from the desktop App. Having the App does not give you the `claude` terminal command; you need to install the CLI separately. You only need it once for this setup.

**Install the CLI.** Open PowerShell and run:

```{r}
#| echo: true
#| eval: false
irm https://claude.ai/install.ps1 | iex
```

**Fix PATH if needed.** The installer may warn that the install location is not in your PATH. If so, run this before closing PowerShell (no need to change anything — it uses your current username automatically):

```{r}
#| echo: true
#| eval: false
[Environment]::SetEnvironmentVariable("PATH", $env:PATH + ";$env:USERPROFILE\.local\bin", "User")
```

Then close and reopen PowerShell.

**Verify.** Run `claude --version`. You should see a version number. If you still get "not recognized," close and reopen PowerShell once more.

### GitHub access

You need access to the `RRC-Associates/claude-rrc-plugins` GitHub repository. No local clone is needed — Claude Code fetches the plugin directly from GitHub.

## Setup

The plugin install commands run inside a Claude Code CLI session — not in PowerShell directly, and not in the desktop App GUI. You only need to do this once. After setup, the skills load automatically in every session including the App GUI.

### Step 1: Open a Claude Code session

In PowerShell, run:

```{r}
#| echo: true
#| eval: false
claude
```

You will be prompted to:

1. **Sign in** — select the option to log in with your Anthropic subscription. A verification code will be displayed; paste it back into the session when prompted.
2. **Pick a mode** — select the default mode to continue.

Once signed in, you'll see the `>` prompt. You're now inside Claude Code.

### Step 2: Add the marketplace

At the `>` prompt, run:

```{r}
#| echo: true
#| eval: false
/plugin marketplace add RRC-Associates/claude-rrc-plugins
```

This registers the RRC plugin repository as a source Claude Code can install from. You only need to do this once.

### Step 3: Install the plugin

Still at the `>` prompt, run:

```{r}
#| echo: true
#| eval: false
/plugin install rrc-R-tools@claude-rrc-plugins
```

Claude Code will show a confirmation screen with the plugin name and description. Select **Install for you (user scope)** and press Enter.

User scope installs the plugin for your account across all projects. Do not choose project scope unless you have a specific reason.

### Step 4: Verify

Close and reopen Claude Code (CLI or App GUI), then start a new session. Ask Claude: "What skills do you have available?" You should see `r-style`, `diagnose-before-fixing`, `quarto-render-sync`, and `quarto-scaffold` in the response.

## Keeping the Plugin Current

Claude Code checks for plugin updates automatically. If you want to force a refresh, run `/plugin update rrc-R-tools@claude-rrc-plugins` at the prompt.

To suggest a change to one of the skills, reach out to Cathy. Changes go into the GitHub repo and propagate to everyone on the next update.
